---
title: "PharmacoGXDataExploration"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

THe goal of this document is to explore the tools available in PharmacoGX with respect to how they can help support CANDLE infrastructure.

```{r load packages}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

if(!require('PharmacoGx')){
  BiocManager::install("PharmacoGx")
  library('PharmacoGx')
}

library(dplyr)
library(tidyr)
all.dsets<-PharmacoGx::availablePSets()


##AFAIK, these are the *new* datasets, 
new.ds<-c('BeatAML','UHNBreast','PRISM','PDTX','FIMM','Tavor','GBM','GRAY')

dsets<-subset(all.dsets,`Dataset Name`%in%new.ds)
print(dsets)
```

Here we have the data sets and need to move forward to download them. This can take a while.

```{r download data}
top5<-dsets$`PSet Name`[1:4]
bottom5<-dsets$`PSet Name`[5:8]

dsetList<-lapply(top5,function(x) PharmacoGx::downloadPSet(x))
names(dsetList)<-new.ds[1:4]

otherDset<-lapply(bottom5,function(x) PharmacoGx::downloadPSet(x))
names(otherDset)<-new.ds[5:8]

fulllist<-c(dsetList,otherDset)

```


## Number of drugs/cell lines in each dataset

Each of the 5 datasets has different cell lines, drug data. Now we want to collect the cell line names, the data available, and number of drugs for the datasets that are new to the system.

```{r data summary}

##such a garbage API with no standardization
molmapping<-list(BeatAML='rnaseq.counts',
                 UHNBreast='Kallisto_0.46.1.rnaseq',PRISM=NA,
                 PDTX='rna',FIMM=NA,Tavor='rnaseq',GBM='rna',GRAY='rna')

drugmapping<-list(BeatAML='aac_recomputed',UHNBreast='aac_recomputed',
                  PRISM='aac_recomputed',PDTX='aac_recomputed',FIMM='',
                  Tavor='',GBM='',GRAY='')

stats<-lapply(new.ds,function(dsetname){
  print(dsetname)
  dset<-fulllist[[dsetname]]
  dvals<-dim(summarizeSensitivityProfiles(dset,'aac_recomputed'))
  
  moltype<-molmapping[[dsetname]]
  if(!is.na(moltype))
    rvals<-dim(summarizeMolecularProfiles(dset,moltype))
  else
    rvals<-c(0,0)
  
  return(list(chems=dvals[1],cellChems=dvals[2],genes=rvals[1],cellRNA=rvals[2],dataset=dsetname))
})

tab<-do.call(rbind,stats)%>%
  as.data.frame()%>%
  tidyr::pivot_longer(c(cellChems,cellRNA),names_to='dataType',values_to='Number cell lines')%>%
  mutate(chems=unlist(chems),`Number cell lines`=unlist(`Number cell lines`),dataset=unlist(dataset))

library(ggplot2)

ggplot(tab,aes(x=dataset,y=`Number cell lines`,fill=dataType))+geom_bar(stat='identity',position='dodge')+ggtitle('Number of cell lines measured')
ggsave('cellLines.png')

ggplot(subset(tab,dataType=='cellChems'),aes(x=dataset,y=`chems`))+geom_bar(stat='identity',position='dodge')+ggtitle('Number of chemicals measured')
ggsave('chems.png')


```


## Drug/cell line overlap

Next up we need to determine what are the unique cell lines and drugs being added to the portfolio.


```{r plot stats}
library(UpSetR)
cells<-lapply(new.ds,function(dsetname){
  dset<-fulllist[[dsetname]]
  cellNames(dset)
  
})
names(cells)<-new.ds

drugs<-lapply(new.ds,function(dsetname){
  dset<-fulllist[[dsetname]]
  drugNames(dset)

})
names(drugs)<-new.ds

png('cellLineOverlap.png')
upset(fromList(cells),nsets=8,mainbar.y.label = 'Cell Lines')
dev.off()

png('drugOverlap.png')
upset(fromList(drugs),nsets=8,mainbar.y.label="Drugs across datasets")
dev.off()
```
Clearly these new cell lines and drug treatments are distinct from each other, but how do they compare to the datasets that have already been collected?

```{r assess prior data}

all.dsets<-PharmacoGx::availablePSets()


other.ds<-subset(all.dsets,!`Dataset Name`%in%new.ds) ##now we have these other datasets

top5<-other.ds$`PSet Name`[1:4]
#bottom5<-other.ds$`PSet Name`[4:6]

dsetList<-lapply(top5,function(x) PharmacoGx::downloadPSet(x))
names(dsetList)<-other.ds[1:4]

#otherDset<-lapply(bottom5,function(x) PharmacoGx::downloadPSet(x))
#names(otherDset)<-other.ds[4:6]

#otherList<-c(dsetList,otherDset)
names(otherList)<-other.ds
orig.cells<-lapply(other.ds,function(dsetname){
  dset<-otherList[[dsetname]]
  cellNames(dset)
  
})
names(orig.cells)<-other.ds

orig.drugs<-lapply(other.ds,function(dsetname){
  dset<-otherList[[dsetname]]
  drugNames(dset)

})
names(orig.drugs)<-other.ds
```
